---
title: "Symcap"
author: "Teegan Innis"
date: "August 31, 2016"
output: html_document
---
--------------------------------------------------------------------------------------
### Project Summary
This project aimed to look at the spatial variability of Symbiodinium clades C and D in the Kane'ohe Bay, O'ahu, Hawai'i population of Montipora capitata. We investigated the distribution of symbionts at scales ranging from location in the bay to location on an individual reef. We also looked at differences among reef types (fringing vs. patch), colony color morph (brown vs. orange) and depth. Heterogeneous mixtures of symbiont clades were considered in the analysis for spatial patterns. By investigating spatial variability of Symbiodinium, we furthered the understanding of stress-response potential in Kane'ohe Bay.


### Library Packages
```{r Library Packages Needed, warning=FALSE, message=FALSE}
setwd("~/Symcap")
library(data.table)
library(devtools)
library(plyr)
library(reshape2)
library(popbio)
library(RgoogleMaps)
library(plotrix)
library(zoo)
```


### Import/Merge Field and qPCR Data
```{r Import Data, warning=FALSE, message=FALSE}
Coral_Data <- read.csv("Coral_Collection.csv")
Coral_Data$Depth..m. <- as.numeric(as.character(Coral_Data$Depth..m.))
source_url("https://raw.githubusercontent.com/jrcunning/steponeR/master/steponeR.R")
Mcap.plates <- list.files(path = "qPCR_data", pattern = "txt$", full.names = T)
Mcap <- steponeR(files = Mcap.plates, delim="\t",
                 target.ratios=c("C.D"), 
                 fluor.norm = list(C=2.26827, D=0), 
                 copy.number=list(C=33, D=3))
Mcap <- Mcap$result
Mcap <- Mcap[grep("+", Mcap$Sample.Name, fixed=T, invert = T), ]
Mcap <- Mcap[grep("NTC", Mcap$Sample.Name, fixed = T, invert = T), ]
Mcap <- Mcap[grep("PCT", Mcap$Sample.Name, fixed = T, invert = T), ]
colnames(Mcap)[which(colnames(Mcap)=="Sample.Name")] <- "Colony"
Mcap$fail <- ifelse(Mcap$C.reps < 2 & Mcap$D.reps < 2, TRUE, FALSE)
fails <- Mcap[Mcap$fail==TRUE, ]
Mcap <- Mcap[which(Mcap$fail==FALSE),]
Mcap$C.D[which(Mcap$C.reps<2)] <- -Inf
Mcap$C.D[which(Mcap$D.reps<2)] <- Inf
Mcap <- Mcap[with(Mcap, order(Colony)), ]
Mcap$propC <- Mcap$C.D / (Mcap$C.D+1)
Mcap$propD <- 1-Mcap$propC
Mcap$propD[which(Mcap$C.D==-Inf)] <-1
Mcap$propC[which(Mcap$C.D==-Inf)] <-0
Mcap$propD[which(Mcap$C.D==Inf)] <-0
Mcap$propC[which(Mcap$C.D==Inf)] <-1
Mcap$Dom <- ifelse(Mcap$propC>Mcap$propD, "C", "D")
Symcap<-merge(Coral_Data, Mcap, by="Colony", all=T)
Symcap <- Symcap[with(Symcap, order(Colony)), ]
Symcap$Mix <- factor(ifelse(Symcap$propC>Symcap$propD, ifelse(Symcap$propD!=0, "CD", "C"), ifelse(Symcap$propD>Symcap$propC, ifelse(Symcap$propC!=0, "DC", "D"), NA)), levels = c("C", "CD", "DC", "D"))
```


### Adjust Depth by Mean Sea Level
```{r Adjust Depth, warning=FALSE, message=FALSE, results='hide'}
JuneTide=read.csv("Station_1612480_tide_ht_20160601-20160630.csv")
JulyTide=read.csv("Station_1612480_tide_ht_20160701-20160731.csv")
Tide<-rbind(JuneTide, JulyTide)
Tide$Time <- as.POSIXct(Tide$TimeUTC, format="%Y-%m-%d %H:%M:%S", tz="UTC")
attributes(Tide$Time)$tzone <- "Pacific/Honolulu"
Symcap$Time2 <- as.POSIXct(paste(as.character(Symcap$Date), as.character(Symcap$Time)), format="%m/%d/%y %H:%M", tz="Pacific/Honolulu")
Symcap$Time=Symcap$Time2
Round6 <- function (times)  {
  x <- as.POSIXlt(times)
  mins <- x$min
  mult <- mins %/% 6
  remain <- mins %% 6
  if(remain > 3L) {
    mult <- mult + 1
  } else {
    x$min <- 6 * mult
  }
  x <- as.POSIXct(x)
  x
}
Symcap$Time.r <- Round6(Symcap$Time)
Tide$Time.r <- Tide$Time
merged<-merge(Symcap, Tide, by="Time.r", all.x=T)
merged$newDepth <- merged$Depth..m.- merged$TideHT
```


### Coral Collection Map
```{r Collection Map}
KB <- c(21.46087401, -157.809907) 
KBMap <- GetMap(center = KB, zoom = 13, maptype = "satellite", SCALE = 2, GRAYSCALE = FALSE)
Latitude=aggregate(Latitude~Reef.ID, data=Symcap, FUN = mean)
Longitude=aggregate(Longitude~Reef.ID, data = Symcap, FUN=mean)
XY<-merge(Latitude, Longitude, by="Reef.ID", all=T)
newcoords <- LatLon2XY.centered(KBMap, XY$Latitude, XY$Longitude, zoom=13)
XY$X <- newcoords$newX
XY$Y <- newcoords$newY
PlotOnStaticMap(KBMap, XY$Latitude, XY$Longitude, col=153, pch=21, bg="#FF7F50", lwd=2)
```


### Dominant Symbiont by Reef Area
```{r Dom~ReefArea}
Symcap$Reef.Area <- ifelse(Symcap$Reef.Area!="Top", yes = "Slope", no = "Top")
results=table(Symcap$Dom, Symcap$Reef.Area)
chisq.test(results)
prop.table(results, margin = 2)
par(mar=c(4, 4, 2, 6))
barplot(prop.table(results, margin = 2), col = c("gray10", "gray100"), xlab = "Reef Area", ylab = "Proportion of Dominant Symbiont")
legend("topright", legend=c("C", "D"), fill=c("gray10", "gray100"), inset = c(-.2, 0), xpd = NA)
```


### Dominant Symbiont by Color Morph
```{r Dom~ColorMorph}
results=table(Symcap$Dom, Symcap$Color.Morph)
chisq.test(results)
prop.table(results, margin = 2)
par(mar=c(4, 4, 2, 6))
barplot(prop.table(results, margin = 2), col = c("gray10", "gray100"), xlab = "Color Morph", ylab = "Proportion of Dominant Symbiont")
legend("topright", legend=c("C", "D"), fill=c("gray10", "gray100"), inset = c(-.2, 0), xpd = NA)
```


### Symbiont Community Composition by Color Morph
```{r Mix~ColorMorph, warning=FALSE}
results=table(Symcap$Mix, Symcap$Color.Morph)
chisq.test(results)
prop.table(results, margin = 2)
par(mar=c(4, 4, 2, 6))
barplot(prop.table(results, margin = 2), col = c("gray10", "gray50", "gray85", "gray100"), xlab = "Color Morph", ylab = "Symbiont Community Composition")
legend("topright", legend=c("C", "CD", "DC", "D"), fill=c("gray10", "gray50", "gray85", "gray100"), inset = c(-.2, 0), xpd = NA)
```


### Symbiont Community Composition by Reef Area
```{r Mix~ReefArea, warning=FALSE}
Symcap$Reef.Area <- ifelse(Symcap$Reef.Area!="Top", yes = "Slope", no = "Top")
results=table(Symcap$Mix, Symcap$Reef.Area)
chisq.test(results)
prop.table(results, margin = 2)
par(mar=c(4, 4, 2, 6))
barplot(prop.table(results, margin = 2), col = c("gray10", "gray50", "gray85", "gray100"), xlab = "Reef Area", ylab = "Symbiont Community Composition")
legend("topright", legend=c("C", "CD", "DC", "D"), fill=c("gray10", "gray50", "gray85", "gray100"), inset = c(-.2, 0), xpd = NA)
```


### Color Morph by Reef Area
```{r ColorMorph~ReefArea}
Symcap$Reef.Area <- ifelse(Symcap$Reef.Area!="Top", yes = "Slope", no = "Top")
results=table(Symcap$Color.Morph, Symcap$Reef.Area)
chisq.test(results)
prop.table(results, margin = 2)
par(mar=c(4, 4, 2, 6))
barplot(prop.table(results, margin = 2), col = c("gray10", "gray100"), xlab = "Reef Area", ylab = "Color Morph")
legend("topright", legend=c("Brown", "Orange"), fill=c("gray10", "gray100"), inset = c(-.2, 0), xpd = NA)
```


### Dominant Symbiont by Depth
```{r Dom~Depth}
merged$Dominant <- ifelse(merged$Dom=="C", 1, 0)
Dom1 <- subset(merged, !is.na(newDepth) & !is.na(Dominant))
results=glm(Dominant~newDepth, family = "binomial", data = merged)
anova(results, test = "Chisq")
logi.hist.plot(Dom1$newDepth, Dom1$Dominant, boxp = FALSE, type = "hist", col="gray", xlabel = "Depth (m)", ylabel = "", ylabel2 = "")
mtext(side = 4, text = "Frequency", line = 3, cex=1)
mtext(side = 4, text = "C                                       D", line = 2, cex = 0.75)
mtext(side = 2, text = "Probability of clade C Symbiont", line = 3, cex = 1)
```


### Color Morph by Depth
```{r ColorMorph~Depth}
merged$Color <- ifelse(merged$Color.Morph=="Orange", 1, 0)
results=glm(Color~newDepth, family = "binomial", data = merged)
anova(results, test = "Chisq")
Color <- subset(merged, !is.na(newDepth) & !is.na(Color))
logi.hist.plot(independ = Color$newDepth, depend = Color$Color, type = "hist", boxp = FALSE, ylabel = "", col="gray", ylabel2 = "", xlabel = "Depth (m)")
mtext(side = 4, text = "Frequency", line = 3, cex=1)
mtext(side = 4, text = "Brown                                Orange", line = 2, cex = 0.75)
mtext(side = 2, text = "Probability of Orange Color Morph", line = 3, cex = 1)
```